---
title: "Cross Validation"
author: "Duzhi Zhao"
date: "11/12/2019"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 12, 
  fig.height = 10,
  out.width = "100%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
set.seed(1)

```

# Cross validation by hand
```{r}
nonlin_df = 
  tibble(
    id = 1:100,
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
  )

nonlin_df %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  theme_bw()
```

## Training and testing
```{r}
train_df = sample_n(nonlin_df, 80)
test_df = anti_join(nonlin_df, train_df, by = "id") 
# "anti_join" gives data in the nonlin_df, but not in train_df (in this case, pull the rest of 20 data points from nonlin_df)

ggplot(train_df, aes(x = x, y = y)) + 
  geom_point() + 
  geom_point(data = test_df, color = "red")

```

## Fit 3 models of varying goodnes
```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = mgcv::gam(y ~ s(x), data = train_df)
wiggly_mod = mgcv::gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df)
```

### Plot linear model
```{r}
train_df %>% 
  add_predictions(linear_mod) %>% 
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

### Plot Smooth model
```{r}
train_df %>% 
  add_predictions(smooth_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

### Plot wiggly model
```{r}
train_df %>% 
  add_predictions(wiggly_mod) %>% 
  ggplot(aes(x = x, y = y)) + geom_point() + 
  geom_line(aes(y = pred), color = "red")
```

## Compute root mean squared errors (rmse) for each model
```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)


rmse(linear_mod, train_df)
rmse(smooth_mod, train_df)
rmse(wiggly_mod, train_df)
```

# Cross validation using 'modelr'
```{r}
cv_df = 
  crossv_mc(nonlin_df, 100) 
```

```{r}
cv_df %>% 
  pull(train) %>% 
  .[[1]] %>% 
  as_tibble

cv_df %>% 
  pull(test) %>% 
  .[[1]] %>% 
  as_tibble

# The above two extracts data from first row of cv_df

cv_df =
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

## Try fitting linear model to all of these
```{r}
cv_results =
  cv_df %>% 
  mutate(
    linear_mods = map(.x = train, ~lm(y ~ x, data = .x)),
    rmse_lin = map2_dbl(.x = linear_mods, .y = test, ~rmse(.x, .y)),
    
    smooth_mods = map(.x = train, ~gam(y ~ s(x), data = .x)),
    rmse_smooth = map2_dbl(.x = smooth_mods, .y = test, ~rmse(.x, .y)),
    
    wiggly_mods  = map(train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x)),
    rmse_wiggly = map2_dbl(wiggly_mods, test, ~rmse(model = .x, data = .y))          
  )
```

## Visualize this
```{r}
cv_results %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()

# Smooth model is the best
```

Example: Child Growth

```{r}
child_growth = read_csv("./data/nepalese_children.csv")

child_growth %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5)
```

```{r}
child_growth =
  child_growth %>% 
  mutate(weight_cp = (weight > 7) * (weight - 7))
```

```{r}
linear_mod    = lm(armc ~ weight, data = child_growth)
# piecewise linear model
pwl_mod    = lm(armc ~ weight + weight_cp, data = child_growth)
smooth_mod = gam(armc ~ s(weight), data = child_growth)
```
Plot the three models to get intuition for goodness of fit
```{r}
child_growth %>% 
  gather_predictions(linear_mod, pwl_mod, smooth_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```

```{r}
cv_df =
  crossv_mc(child_growth, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))
```

Mutate + map & map2 to fit models to training data and obtain corresponding RMSEs for the testing data.
```{r}
cv_df = 
  cv_df %>% 
  mutate(linear_mod  = map(train, ~lm(armc ~ weight, data = .x)),
         pwl_mod     = map(train, ~lm(armc ~ weight + weight_cp, data = .x)),
         smooth_mod  = map(train, ~gam(armc ~ s(weight), data = as_tibble(.x)))) %>% 
  mutate(rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
         rmse_pwl    = map2_dbl(pwl_mod, test, ~rmse(model = .x, data = .y)),
         rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)))
```

```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

